Measuring Subgroup Preferences in Conjoint Experiments

### 2019-05-24 manuscript.Rnw
final PA submission

\documentclass[a4paper,12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{mathpazo}
%\usepackage{times}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage[normalem]{ulem}
\usepackage{sectsty}
\subsectionfont{\normalfont\large\underline}
\usepackage{floatrow}
\floatsetup[figure]{capposition=top}
\floatsetup[table]{capposition=top}

\title{Measuring Subgroup Preferences in Conjoint Experiments\footnote{We thank Benjamin Lauderdale, Jamie Druckman, Yusaku Horiuchi, the editor, and anonymous reviewers for feedback on this manuscript.}}
\author{Thomas J. Leeper, Sara B. Hobolt, and James Tilley}

@@

<<hidden, eval=FALSE, echo=FALSE, results="hide">>=
# setwd("c:/users/thomas/dropbox/brexit identity/papers/subgroupanalysis")
# knitr::knit2pdf("manuscript.Rnw", quiet = TRUE)
@

<<knitr_options, cache=FALSE, echo=FALSE, results="hide">>=
knitr::opts_chunk$set(cache = TRUE, echo=FALSE, results="hold", warning=FALSE, error=FALSE, message=FALSE)
options(scipen = 10)
@

@@

In what follows, we demonstrate the challenges of conjoint analysis and remind readers of how reference category choice for profile features creates problems for comparing conditional AMCEs across respondent subgroups. We show how the use of an arbitrary reference category means the size, direction, and statistical significance of differences-in-AMCEs have little relationship to the underlying degree of favorability of the subgroups toward profiles with particular features. Reference category choices can make similar preferences look dissimilar and dissimilar preferences look similar. We demonstrate this with examples drawn from the published political science literature (namely experiments by \citealt{HainmuellerHopkinsYamamoto2014, BechtelScheve2013, TeeleKallaRosenbluth2018}). The paper then provides suggestions for improved conjoint reporting and interpretation based around two quantities of interest drawn from the factorial experimentation literature: (a) unadjusted marginal means, a quantity measuring favorability toward a given feature, and (b) an omnibus F-test, measuring differences therein. Software for the R programming language to support our findings --- and that can be used to examine sensitivity of conjoint analysis to reference category selection, calculate AMCEs and marginal means, perform subgroup analyses, and test for subgroup differences in any conjoint experiment --- is demonstrated throughout. We conclude with advice for best practices in the analysis and presentation of conjoint results.

@@

<<tkr_replication, dependson=c("data_tkr"), fig.height=4.5, fig.width=5, fig.cap="Replication of Results for `Candidate Sex' Feature from Teele et al. (2018) Candidate Experiment using Full Sample and Subgroup AMCEs and MMs", fig.show="hold", message=FALSE>>=
a <- plot(
  subset(cj(
   tkr, 
   winner ~ feature_sex,
   id = ~ responseid,
   estimate = "amce",
  ), feature == "Candidate Sex"), 
  xlab = NULL, feature_headers = FALSE, vline = 0, xlim = c(-0.11, 0.11), size = 2
) + 
  ggtitle(NULL, "Full Sample AMCEs") + 
  scale_colour_manual("Feature", limits = "Candidate Sex", values = "black") +
  theme(
  	legend.position = "none",
  	panel.border = element_blank(),
  	axis.line.x = element_blank(),
	axis.line.y = element_blank()
  )

@@

c <- plot(
  subset(cj(
   tkr, 
   winner ~ feature_sex,
   id = ~ responseid,
   estimate = "amce",
   by = ~ PartyID
  ), feature == "Candidate Sex"), 
  xlab = NULL, feature_headers = FALSE, vline = 0, group = "PartyID", xlim = c(-0.11, 0.11), size = 2
) + 
 annotate("text", y = 1.65, x = -0.04, label = "Republicans", size = 3) + 
 annotate("text", y = 1.3, x = 0.05, label = "Democrats", size = 3) + 
 ggtitle(NULL, "Subgroup AMCEs") + 
 scale_colour_manual(values=rep("black", 2)) +
 aes(shape = PartyID) + 
 scale_shape_manual("PartyID", limits = c("Democrat", "Republican"), values = c(15,17)) +
 theme(
   legend.position = "none",
   panel.border = element_blank(),
   axis.line.x = element_blank(),
   axis.line.y = element_blank()
 )

d <- plot(
  subset(cj(
   tkr, 
   winner ~ feature_sex,
   id = ~ responseid,
   estimate = "mm",
   by = ~ PartyID
  ), feature == "Candidate Sex"),
  xlab = NULL, feature_headers = FALSE, vline = 0.5, group = "PartyID", xlim = c(0.39,0.61), size = 2
) + 
 ggtitle(NULL, "Subgroup MMs") + 
 scale_colour_manual(values=rep("black", 2)) +
 aes(shape = PartyID) + 
 scale_shape_manual("PartyID", limits = c("Democrat", "Republican"), values = c(15,17)) +
 theme(
  legend.position = "none",
  panel.border = element_blank(),
  axis.line.x = element_blank(),
  axis.line.y = element_blank()
 )

@@

Panel A gives the impression that there are significant differences in preferences between high and low ethnocentrism respondents toward immigrants from different countries of origin, with different careers, and with different educational attainments because the reference category choice cascades the difference in reference category favorability into AMCEs for all other feature levels. By contrast, Panel B gives the impression that these differences are negligible. The experimental data and analytic approach in the two portrayals is identical; the only difference is the choice of reference category. Given what we have shown about the relationship between differences in conditional AMCEs and differences in conditional marginal means, Panel B is the more truthful visualization \citep{Cairo2016}. The differences between subgroup AMCEs there more accurately convey differences in underlying preferences because the reference categories used in Panel B are the most similar between the two groups.

@@

\appendix
\tableofcontents

@@

c <- plot(
  subset(cj(
   tkr, 
   winner ~ feature_sex + feature_experience + feature_marital + feature_job + feature_children + feature_age,
   id = ~ responseid,
   estimate = "amce",
   by = ~ PartyID
  ), feature == "Political Experience"), 
  xlab = NULL, feature_headers = FALSE, vline = 0, group = "PartyID", xlim = c(-0.21, 0.21), size = 2
) + 
 annotate("text", y = 1.65, x = -0.04, label = "Republicans", size = 3) + 
 annotate("text", y = 1.3, x = 0.05, label = "Democrats", size = 3) + 
 ggtitle(NULL, "Subgroup AMCEs") + 
 scale_colour_manual(values=rep("black", 2)) +
 aes(shape = PartyID) + 
 scale_shape_manual("PartyID", limits = c("Democrat", "Republican"), values = c(15,17)) +
 theme(
   legend.position = "none",
   panel.border = element_blank(),
   axis.line.x = element_blank(),
   axis.line.y = element_blank()
 )

d <- plot(
  subset(cj(
   tkr, 
   winner ~ feature_sex + feature_experience + feature_marital + feature_job + feature_children + feature_age,
   id = ~ responseid,
   estimate = "mm",
   by = ~ PartyID
  ), feature == "Political Experience"),
  xlab = NULL, feature_headers = FALSE, vline = 0.5, group = "PartyID", xlim = c(0.29,0.71), size = 2
) + 
 ggtitle(NULL, "Subgroup MMs") + 
 scale_colour_manual(values=rep("black", 2)) +
 aes(shape = PartyID) + 
 scale_shape_manual("PartyID", limits = c("Democrat", "Republican"), values = c(15,17)) +
 theme(
  legend.position = "none",
  panel.border = element_blank(),
  axis.line.x = element_blank(),
  axis.line.y = element_blank()
 )

>>>>

\documentclass[a4paper,12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{mathpazo}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage[normalem]{ulem}
\usepackage{sectsty}
\subsectionfont{\normalfont\large\underline}
\usepackage{floatrow}
\usepackage[hidelinks]{hyperref}
\floatsetup[figure]{capposition=top}
\floatsetup[table]{capposition=top}

\title{Measuring Subgroup Preferences in Conjoint Experiments\footnote{We thank Benjamin Lauderdale, Jamie Druckman, Yusaku Horiuchi, the editor, and anonymous reviewers for feedback on this manuscript. Replication data and code for this article are available from the \textit{Political Analysis} Dataverse: \url{https://doi.org/10.7910/DVN/ARHZU4}. This work was funded, in part, by the United Kingdom Economic and Social Research Council (Grant ES/R000573/1).}
}
\author{Thomas J. Leeper, Sara B. Hobolt, and James Tilley}

@@

<<knitr_options, cache=FALSE, echo=FALSE, results="hide">>=
knitr::opts_chunk$set(
  cache = TRUE,
  echo=FALSE,
  results="hold",
  warning=FALSE,
  error=FALSE,
  message=FALSE
)
options(scipen = 10)
@

@@

In what follows, we demonstrate the challenges of conjoint analysis and remind readers of how reference category choice for profile features creates problems for comparing conditional AMCEs across respondent subgroups. We show how the use of an arbitrary reference category means the size, direction, and statistical significance of differences-in-AMCEs have little relationship to the underlying degree of favorability of the subgroups toward profiles with particular features. Reference category choices can make similar preferences look dissimilar and dissimilar preferences look similar. We demonstrate this with examples drawn from the published political science literature (namely experiments by \citealt{HainmuellerHopkinsYamamoto2014, BechtelScheve2013, TeeleKallaRosenbluth2018}). The paper then provides suggestions for improved conjoint reporting and interpretation based around two quantities of interest drawn from the factorial experimentation literature: (a) unadjusted marginal means, a quantity measuring favorability toward a given feature, and (b) an omnibus F-test, measuring differences therein. Software for the R programming language to support our findings --- and that can be used to examine sensitivity of conjoint analysis to reference category selection, calculate AMCEs and marginal means, perform subgroup analyses, and test for subgroup differences in any conjoint experiment \citep{Leeper2018cregg} --- is demonstrated throughout using example data \citep{LeeperHoboltTilley2019}. We conclude with advice for best practices in the analysis and presentation of conjoint results.

@@

<<tkr_replication, dependson=c("data_tkr"), fig.height=4.5, fig.width=5, fig.cap="Replication of Results for `Candidate Sex' Feature from Teele et al. (2018) Candidate Experiment using Full Sample AMCEs and MMs and Subgroup AMCEs and MMs for Democrats and Republicans", fig.show="hold", message=FALSE>>=
a <- plot(
  subset(cj(
   tkr, 
   winner ~ feature_sex,
   id = ~ responseid,
   estimate = "amce",
  ), feature == "Candidate Sex"), 
  xlab = NULL, feature_headers = FALSE, vline = 0, xlim = c(-0.11, 0.11), size = 2
) + 
  ggtitle(NULL, "Full Sample AMCEs") + 
  scale_colour_manual("Feature", limits = "Candidate Sex", values = "black") +
  theme(
  	legend.position = "none",
  	panel.border = element_blank(),
  	axis.line.x = element_blank(),
	axis.line.y = element_blank()
  )

@@

c <- plot(
  subset(cj(
   tkr, 
   winner ~ feature_sex,
   id = ~ responseid,
   estimate = "amce",
   by = ~ PartyID
  ), feature == "Candidate Sex"), 
  xlab = NULL, feature_headers = FALSE, vline = 0, group = "PartyID", xlim = c(-0.11, 0.11), size = 2
) + 
 annotate("text", y = 1.7, x = -0.04, label = "Republicans", size = 3) + 
 annotate("text", y = 1.35, x = 0.05, label = "Democrats", size = 3) + 
 ggtitle(NULL, "Subgroup AMCEs") + 
 scale_colour_manual(values=rep("black", 2)) +
 aes(shape = PartyID) + 
 scale_shape_manual("PartyID", limits = c("Democrat", "Republican"), values = c(15,17)) +
 theme(
   legend.position = "none",
   panel.border = element_blank(),
   axis.line.x = element_blank(),
   axis.line.y = element_blank()
 )

d <- plot(
  subset(cj(
   tkr, 
   winner ~ feature_sex,
   id = ~ responseid,
   estimate = "mm",
   by = ~ PartyID
  ), feature == "Candidate Sex"),
  xlab = NULL, feature_headers = FALSE, vline = 0.5, group = "PartyID", xlim = c(0.39,0.61), size = 2
) + 
 annotate("text", y = 1.8, x = 0.48, label = "Rep.", size = 3) + 
 annotate("text", y = 0.8, x = 0.52, label = "Rep.", size = 3) + 
 annotate("text", y = 2.2, x = 0.55, label = "Dem.", size = 3) + 
 annotate("text", y = 1.2, x = 0.45, label = "Dem.", size = 3) + 
 ggtitle(NULL, "Subgroup MMs") + 
 scale_colour_manual(values=rep("black", 2)) +
 aes(shape = PartyID) + 
 scale_shape_manual("PartyID", limits = c("Democrat", "Republican"), values = c(15,17)) +
 theme(
  legend.position = "none",
  panel.border = element_blank(),
  axis.line.x = element_blank(),
  axis.line.y = element_blank()
 )

@@

Panel A gives the impression that there are significant differences in preferences between high and low ethnocentrism respondents toward immigrants from different countries of origin, with different careers, and with different educational attainments because the reference category choice cascades the difference in reference category favorability into AMCEs for all other feature levels. By contrast, Panel B gives the impression that these differences are negligible. The experimental data and analytic approach in the two portrayals is identical; the only difference is the choice of reference category. Given what we have shown about the relationship between differences in conditional AMCEs and differences in conditional marginal means, Panel B is a more ``truthful'' visualization, which \citet{Cairo2016} uses to mean avoidance of self-deception in the presentation of data, and a more ``functional'' visualization, by which \citeauthor{Cairo2016} means choosing graphics based on how they will be interpreted by the visualization's consumers. The differences between subgroup AMCEs there more accurately convey differences in underlying preferences because the reference categories used in Panel B are the most similar between the two groups.

@@

\appendix
\setcounter{page}{1}

\noindent {\large ``Measuring Subgroup Preferences in Conjoint Experiments: Supplemental Information''}

\vspace{0.5em}

\noindent Thomas J. Leeper, Sara B. Hobolt, and James Tilley\\
\noindent 2019-05-24

\renewcommand{\contentsname}{}
\vspace{-2em}
\tableofcontents

@@

c <- plot(
  subset(cj(
   tkr, 
   winner ~ feature_sex + feature_experience + feature_marital + feature_job + feature_children + feature_age,
   id = ~ responseid,
   estimate = "amce",
   by = ~ PartyID
  ), feature == "Political Experience"), 
  xlab = NULL, feature_headers = FALSE, vline = 0, group = "PartyID", xlim = c(-0.21, 0.21), size = 2
) + 
 annotate("text", y = 1.65, x = -0.04, label = "Republicans", size = 3) + 
 annotate("text", y = 1.3, x = 0.05, label = "Democrats", size = 3) + 
 ggtitle(NULL, "Subgroup AMCEs") + 
 scale_colour_manual(values=rep("black", 2)) +
 aes(shape = PartyID) + 
 scale_shape_manual("PartyID", limits = c("Democrat", "Republican"), values = c(15,17)) +
 theme(
   legend.position = c(0.1,0.8),
   legend.title=element_blank(),
   panel.border = element_blank(),
   axis.line.x = element_blank(),
   axis.line.y = element_blank()
 )

d <- plot(
  subset(cj(
   tkr, 
   winner ~ feature_sex + feature_experience + feature_marital + feature_job + feature_children + feature_age,
   id = ~ responseid,
   estimate = "mm",
   by = ~ PartyID
  ), feature == "Political Experience"),
  xlab = NULL, feature_headers = FALSE, vline = 0.5, group = "PartyID", xlim = c(0.29,0.71), size = 2
) + 
 ggtitle(NULL, "Subgroup MMs") + 
 scale_colour_manual(values=rep("black", 2)) +
 aes(shape = PartyID) + 
 scale_shape_manual("PartyID", limits = c("Democrat", "Republican"), values = c(15,17)) +
 theme(
  legend.position = c(0.1,0.8),
  legend.title=element_blank(),
  panel.border = element_blank(),
  axis.line.x = element_blank(),
  axis.line.y = element_blank()
 )
